To log or not to log

### 2022-08-24 analysis/Analysis ECDC forecasts.Rmd
update plots

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)

@@

# format plot scales
scale_fn <- function(x) {
  ifelse(x >= 1000000, 
         paste0(x / 1000000, "m"), 
         ifelse(x >= 5000,
                paste0(x / 1000, "k"),
                x))
}

@@

## =============================================================================
## FIGURE 4
## Mean vs. sd of observed values

@@

@@

plot_mean_scores <- plot_df |>
  ggplot(aes(y = interval_score,
             x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(scale ~ target_type, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
  labs(y = "Interval score", x = "Location")

plot_means_obs / 
  plot_mean_scores + 
  plot_layout(heights = c(1, 2))

ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 7)

@@

Figure 5? 
that shows Variance vs. mean of the observations. Makes the point that the
data looks neg. binomial. 
```{r}
# plot with mean vs. variance --------------------------------------------------
plot_var_mean <- function(hub_data, log = FALSE) {
  df <- hub_data |>
    dplyr::select(location, target_type, target_end_date, true_value) |>
    unique() |>
    group_by(location, target_type)

@@

    f <- function(x) {log(x + x^2 + 1)}
  }

  df |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  labs(y = "Variance observed value", x = "Mean observed value") + 
  geom_function(fun = f, linetype = "dashed", size = 0.5) + 
  theme_scoringutils() + 
  geom_smooth(alpha = 0.2, size = 0.2)
}

@@

  total_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(total = sum(true_value))

@@

p3 <- wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10")

p1 / p2 / p3 + 
  plot_layout(heights = c(1, 1, 2))

ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 6)

# ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 4)

# ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 2.5)
```

@@

scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")

@@

```{r}
v1 <- c("a","b","c","d","e","f","g")
v2 <- c("b","x","e","c","z","d","a")

o <- 0.05
DF <- data.frame(x = c(rep(1, length(v1)), rep(2, length(v2))),
                 x1 = c(rep(1 + o, length(v1)), rep(2 - o, length(v2))),
                 y = c(rev(seq_along(v1)), rev(seq_along(v2))),
                 g = c(v1, v2))

library(ggplot2)
library(grid)
ggplot(DF, aes(x=x, y=y, group=g, label=g)) +
  geom_path(aes(x=x1), arrow = arrow(length = unit(0.02,"npc")), 
            size=1, color="green") +
  geom_text(size=10) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())
```

## Look at pairwise comparisons

```{r}
summarised_pairwise <- scores |>
  filter(target_type == "Cases") |>
  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)

@@

plot_distr_scores <- scores |>
  filter(horizon == 2) |>
  mutate(model = factor(model, levels = ranking_natural)) |>
  ggplot(aes(y = model, x = interval_score)) +
  ggdist::stat_pointinterval() +
  facet_wrap(~ scale, scale = "free_x") +
  theme_scoringutils() +
  scale_x_log10(labels = label_fn) +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  labs(y = "Model", x = "WIS")

@@

plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores +
  plot_layout(widths = c(2, 1, 2)) &
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

@@

#Issues: this currently ignores that some models have only made forecasts for one location. They should usually be excluded, but maybe not necessary here as we're really more comparing the results for natural and log rather than the models against each other.
```

@@

@@

p_cor_ranks <- correlation_ranks |>
  ggplot(aes(x = horizon, y = cor, 
             fill = `Target type`, color = `Target type`)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Corr. ranks single forecasts", x = "Forecast horizon")

@@

@@

Example comparisons of forecasts on the log and the natural scale

@@

plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases") {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

@@

  plot_natural  <- filtered_hub |>  
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    theme(legend.position = "bottom")

  plot_log <- filtered_hub |>
    mutate(prediction = log(prediction + 1), 
           true_value = log(true_value + 1)) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    theme(legend.position = "bottom")

  scores_natural <- filtered_scores |>
    filter(scale == "natural") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    scale_x_continuous(labels = label_fn)

  scores_log <- filtered_scores |>
    filter(scale == "log") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE)

  (plot_natural + plot_log) /
    (scores_natural + scores_log) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}

ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")

epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")

BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")

ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)
ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 9, height = 5, plot = BSLCoV)

@@

>>>>

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)
library(grid)

@@

# format plot scales
scale_fn <- function(x) {
  ifelse(x >= 1000000, 
         paste0(x / 1000000, "m"), 
         ifelse(x >= 1000,
                paste0(x / 1000, "k"),
                x))
}

@@

## =============================================================================
## FIGURE 4
## Mean vs. sd of observed values + boxplots for distribution across target type + location

@@

box_plot_obs <- hub_data |>
  mutate(true_value = pmax(true_value, 1)) |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  ggplot(aes(y = true_value, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Observations", x = "Target type")

@@

plot_mean_scores <- plot_df |>
  ggplot(aes(y = interval_score,
             x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(scale ~ target_type, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
  labs(y = "Mean interval score", x = "Location")

box_plot_scores <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free", nrow = 2) + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")

box_plot_horizon <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon <= 4) |>
  group_by(forecast_date, model, location, target_type, scale) |>
  mutate(interval_score = interval_score / interval_score[horizon == 1]) |>
  ggplot(aes(y = interval_score,
             x = as.factor(horizon), fill = target_type)) + 
  geom_hline(aes(yintercept = 1), linetype = "dashed", color = "grey40") +
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) +
  geom_boxplot(alpha = 0.5, position = "dodge2") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, nrow = 1) + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(y = "Rel. increase in WIS", x = "Forecast horizon (weeks)")

layout <- "
AB
CD
CD
EE"

plot_means_obs  + box_plot_obs + plot_mean_scores + box_plot_scores + box_plot_horizon +
  plot_layout(heights = c(1, 2, 1), 
              widths = c(3, 1)) +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")

ggsave("output/figures/HUB-mean-obs-location.png", width = 10, height = 10)

@@

Figure
Currently discontinuoued and just there to discuss it. 
that shows Variance vs. mean of the observations. Makes the point that the
data looks neg. binomial. 
```{r}
# plot with mean vs. variance --------------------------------------------------
plot_var_mean <- function(hub_data, log = FALSE) {
  df <- hub_data |>
    mutate("x" = true_value) |>
    dplyr::select(location, target_type, target_end_date, true_value, x) |>
    unique() |>
    group_by(location, target_type)

@@

    f <- function(x) {log(log(x) + log(x)^2 + 1)}
  }

  out <- df |>
  summarise(mean_value = mean(x), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  labs(y = "Var observations", x = "Mean observations") + 
  # geom_function(fun = f, linetype = "dashed", size = 0.5) +
  theme_scoringutils() + 
  geom_smooth(alpha = 0.2, size = 0.2)

  if (log) {
    out <- out +
      labs(y = "Var log(observations)", x = "Mean observations")
  }

  return(out)
}

@@

  total_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(total = mean(true_value))

@@

scatter_wis_obs <- wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10")

p1 / p2 / scatter_wis_obs + 
  plot_layout(heights = c(1, 1, 2))

ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 7)
```

## Figure 5 - Correlations between everything
```{r}
df <- scores |>
  filter(horizon < 5) |>
  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction, type_and_scale)) |>
  pivot_wider(names_from = scale, values_from = interval_score) |> 
  group_by(horizon, target_type, forecast_date, target_end_date, location) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman")) |>
  group_by(horizon, target_type) |>
  summarise(cor = mean(cor, na.rm = TRUE), 
            spearman = mean(cor, na.rm = TRUE)) |>
  rename(`Target type` = target_type)

p_cor_scores <- df |>
  ggplot(aes(x = horizon, y = cor, 
             fill = `Target type`, color = `Target type`)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Avg. Spearman cor(scores)", x = "Forecast horizon")

(scatter_wis_obs) / (p_cor_scores + p_cor_skill) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1, 1), heights = c(2, 1)) &
  theme(legend.position = "bottom")

ggsave("output/figures/HUB-correlations.png", width = 7, height = 6)

@@

@@

## Figure 6 - Look at pairwise comparisons and changes in rankings

```{r}
summarised_pairwise <- scores |>
  filter(target_type == "Cases", 
         horizon == 2) |>
  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)

@@

# plot_distr_scores <- scores |>
#   filter(horizon == 2) |>
#   mutate(model = factor(model, levels = ranking_natural)) |>
#   ggplot(aes(y = model, x = interval_score)) +
#   ggdist::stat_pointinterval() +
#   facet_wrap(~ scale, scale = "free_x") +
#   theme_scoringutils() +
#   scale_x_log10(labels = label_fn) +
#   theme(axis.title.y = element_blank(), 
#         axis.text.y = element_blank(), 
#         axis.ticks.y = element_blank(), 
#         axis.line.y = element_blank()) +
#   labs(y = "Model", x = "WIS")

@@

plot_ranking_change <- summarised_pairwise |>
  select(model, scale, relative_skill) |>
  group_by(scale) |>
  mutate(rank = rank(-relative_skill), 
         x = ifelse(scale == "natural", 0.4, 2.8), 
         x_arrow = ifelse(scale == "natural", 1.7, 2.7)) |>
  group_by(model) |>
  mutate(color = ifelse(diff(relative_skill) > 0, "deteriorated", "not deteriorated")) |>
  ggplot(aes(y = rank, x = x, group = model, label = model)) +
    geom_path(aes(x=x_arrow, color = color), arrow = arrow(length = unit(0.02,"npc")), 
            size=1, show.legend = FALSE) +
  geom_text(size=3.5, hjust = 0) +
  coord_cartesian(xlim = c(0.5, 4)) +
  theme_scoringutils() +
  theme(axis.title = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(), 
        legend.position="none")

layout <- "
AAA
BCD"

plot_ranking_change + plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores 
  plot_layout(design = layout) +
  plot_layout(heights = c(1, 2)) +
  plot_layout(widths = c(2, 1, 2)) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

@@

# summarised_pairwise |>
#   select(model, scale, relative_skill) |>
#   pivot_wider(names_from = scale, values_from = relative_skill) |>
#   ggplot(aes(y = log, x = natural)) +
#   geom_point() + 
#   # geom_abline(slope = 1) +
#   theme_scoringutils()

@@

# we could also do spearman per target and then plot an average of the spearman correlation.

@@

p_cor_ranks <- correlation_ranks |>
  ggplot(aes(x = horizon, y = cor, 
             fill = `Target type`, color = `Target type`)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Corr. ranks per forecast", x = "Forecast horizon")

@@

combined <- scores[, c("model", "location", "target_end_date", "scale", "target_type", "interval_score", "horizon")] |>
  full_join(hub_data[, c("location", "target_end_date", "target_type", "model", "true_value")]) |>
  mutate(true_value = ifelse(scale == "natural", true_value, log(true_value + 1)))

combined |>
  filter(horizon <= 4) |>
  rename(`Target type` = target_type) |>
  group_by(horizon, scale, `Target type`) |>
  summarise(cor = cor(interval_score, true_value)) |>
    ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  facet_wrap(~ scale) +
  expand_limits(ylim = c(0, 1)) +
  labs(y = "Correlation score and observation", x = "Forecast horizon")

@@

## Figure 7 - Example comparisons of forecasts on the log and the natural scale

@@

plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases", 
                            remove_x_text = TRUE) {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

@@

  plot_natural  <- filtered_hub |>  
    mutate(id = paste(type, "- natural")) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    facet_wrap(~ id) +
    theme(legend.position = "bottom", 
          axis.text.x = element_blank(), 
          axis.title.x = element_blank()) +
    labs(y = "Predictions")

  plot_log <- filtered_hub |>
    mutate(prediction = log(prediction + 1), 
           true_value = log(true_value + 1)) |>
    mutate(id = paste(type, "- log")) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    facet_wrap(~ id) +
    theme(legend.position = "bottom", 
          axis.text.x = element_blank(), 
          axis.title.x = element_blank()) +
    labs(y = "Predictions")

  scores_natural <- filtered_scores |>
    filter(scale == "natural") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    scale_x_continuous(labels = label_fn) +
    labs(y = "Target end date")

  scores_log <- filtered_scores |>
    filter(scale == "log") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    labs(y = "Target end date")

  if (remove_x_text) {
    scores_natural <- scores_natural +
      theme(axis.text.x = element_blank(), 
            axis.title.x = element_blank())
    # scores_log <- scores_log +
    #   theme(axis.text.x = element_blank(), 
    #         axis.title.x = element_blank())
  }

  list(plot_natural, scores_natural, plot_log, scores_log )
}

# ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
# epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
# BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
# ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
# ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)

ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                  scores, "EuroCOVIDhub-ensemble", 
                                  remove_x_text = TRUE, 
                                  type = "Cases")

ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                   scores, "EuroCOVIDhub-ensemble", 
                                   type = "Deaths")

put_plot_together <- function(plot_cases, 
                              plot_deaths) {
  layout <- "
AE
BF
CG
DH"

  plot_cases[[1]] + plot_cases[[2]] + plot_cases[[3]] +
    plot_cases[[4]] + plot_deaths[[1]] +
    plot_deaths[[2]] + plot_deaths[[3]] + plot_deaths[[4]] +
    plot_annotation(tag_levels = "A") +
    plot_layout(guides = "collect", 
                design = layout) &
    theme(legend.position = "bottom")
}

put_plot_together(ensemble_cases, ensemble_deaths)
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)

# baseline <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "EuroCOVIDhub-baseline") 
# epinow <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "epiforecasts-EpiNow2") 
# 
# ensemble / epinow +
#   plot_layout(guides = "collect", 
#               heights = c(1, 1, 2)) +
#   theme(legend.position = "bottom")
# 
# ggsave(filename = "output/figures/HUB-model-comparison.png", width = 9, height = 7)

@@

```{r}
scores |>
  summarise_scores(c("model", "target_type", "scale")) |>
  summarise_scores(c("model", "target_type", "scale"), fun = signif, digits = 2) |>
  plot_score_table(by = c("target_type", "scale")) +
  facet_wrap(target_type ~ scale)
```

split_data <- split(scores, by = c("type_and_scale"))
cors <- lapply(split_data, FUN = correlation, 
               metrics = c("interval_score", 
                            "bias", "overprediction", 
                            "underprediction",
                            "dispersion"))

# scoringutils dev stuff
# split_data <- split(scores, by = c("target_type", "scale"))
# list_name_as_col <- function(list, by = NULL) {
#   names <- names(list)
#   print(names(list_element))
# }

plots <- lapply(cors, plot_correlation)

(plots[[1]] + plots[[2]]) /
  (plots[[3]] + plots[[4]])

correlation
```

### 2022-08-24 main.tex
update plots

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
    \caption{Observed mean against observed variance of observations in different locations across time. The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value. }
    \label{fig:HUB-variance-mean}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
    \caption{Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
    \label{fig:HUB-average-scores}
\end{figure}

SECTION STRUCUTRE / IMPORTANT THINGS
\begin{itemize}
\item X variation gets smaller / more even
\item X relationship with country size changes
\item X inndividual rankings stay similar
\item relative skill scores change a lot
\item deep dive into what happens for individual models
\item scores increase across forecast horizons
\end{itemize}

@@

\paragraph{Mean / sd of cases and WIS values}
Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}). The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B). Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
%Could fit a model var = alpha * x2 + beta *x

On the natural scale, scores vary considerably across targets (see Fig \ref{}) and locations (see Fig \ref{}) and correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B). On the log scale, scores are more evenly distributed between targets (see Fig \ref{}) and locations (see Fig \ref{fig:HUB-mean-locations}. The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative. Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented. 
%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-obs-location.png}
    \caption{Scores and average observed values across 1ocations. A, B: Mean of observed cases and deaths across all time points for different locations. C, D: XXX WIS for forecasts made by the Hub ensemble. D, E: Average WIS for log-transformed forecasts. }
    \label{fig:HUB-mean-locations}
\end{figure}

\paragraph{individual rankings}
For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}). When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably. At least for cases, the effect is stronger with increasing forecast horizon.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-cors.png}
    \caption{Correlation between scores on the natural and on the log scale. A: Correlation between relative ranks (the rank in terms of WIS divided by the number of models) for individual forecasts, separated by forecast horizon. B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon. SHOULD THIS BE SPEARMAN?}
    \label{fig:HUB-cors}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-horizon.png}
    \caption{Weighted interval scores across forecast horizons}
    \label{fig:HUB-scores-horizon}
\end{figure}

@@

\paragraph{pairwise comparisons}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
    \caption{Changes in relative skill and the distribution of scores. MAYBE THIS SHOULD ONLY BE CASES?}
    \label{fig:HUB-pairwise}
\end{figure}

% Correlation between pairwise comparisons is 0.4414717.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-ICM.png}
    \caption{Top: 2-week-ahead predictions. Bottom: scores }
    \label{fig:HUB-model-ICM}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-epiBATS.png}
    \caption{Top: 2-week-ahead predictions. Bottom: scores }
    \label{fig:HUB-model-epiBATS}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-BSLCoV.png}
    \caption{Top: 2-week-ahead predictions. Bottom: scores }
    \label{fig:HUB-model-BSLCoV}
\end{figure}

>>>>

SECTION STRUCUTRE / IMPORTANT THINGS
\begin{itemize}
\item X variation gets smaller / more even
\item X relationship with country size changes
\item X inndividual rankings stay similar
\item X relative skill scores change a lot
\item deep dive into what happens for individual models
\item scores increase across forecast horizons
\end{itemize}

@@

\paragraph{Variation of scores gets smaller}
The average number of observed cases and deaths varies considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B). On the natural scale, scores also vary considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C). On the log scale, scores are more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}.

\paragraph{Correlation WIS - obs, location relationship}
On the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B). The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative. Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented.

For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}). When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably. At least for cases, the effect is stronger with increasing forecast horizon.

For relative skill scores, however, rankings do change a lot. The correlation between skill scores for two week ahead forecasts on the log and natural scale is XX. The correlation is stronger for deaths than for cases and tends to be lower for longer horizons. This might be because outliers are more likely for case forecasts than for death forecasts and also are more likely with increasing forecast horizon.

\paragraph{Deep dive}
Differences are likely driven by
- on average, under-prediction plays a larger role (in relative terms), at least for cases. 
- confidence intervals that go very low can produce artifacts --> baseline plot, not shown. 
- natural scale punishes missing the peak very harshly. Log scale also punishes when you miss a new upswing. Arguably this is very good and potentially even closer to what we want.

% Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
%Could fit a model var = alpha * x2 + beta *x
%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-obs-location.png}
    \caption{Scores and average observed values across 1ocations. A, Mean of observed cases and deaths across all time points for different locations. B: XXX WIS for forecasts made by the Hub ensemble. C: Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right).}
    \label{fig:HUB-mean-locations}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-correlations.png}
    \caption{Correlation between scores on the natural and on the log scale. A: Correlation between relative ranks (the rank in terms of WIS divided by the number of models) for individual forecasts, separated by forecast horizon. B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon. SHOULD THIS BE SPEARMAN?}
    \label{fig:HUB-cors}
\end{figure}

@@

\paragraph{pairwise comparisons}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
    \caption{Changes in relative skill and the distribution of scores.Based on case forecasts only.}
    \label{fig:HUB-pairwise}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-comparison.png}
    \caption{Top: 2-week-ahead predictions. Bottom: scores }
    \label{fig:HUB-model-comparison-ensemble}
\end{figure}
